---
title: "Text analysis II: Commending the tm package"
subtitle: "Introduction to R for Social Sciences"
author: 
- "Josef Ginnerskov"
# Uncomment the line below for adding more authors
#- "Another Author"
institute: "Department of Sociology"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    self_contained: true # This allows to use the html file without copying all files.
    lib_dir: libs
    css: ["./resources/css/uppsala.css", "./resources/css/uppsala-fonts.css"]
    includes:
      in_header: "./libs/partials/header.html"
#      after_body: "./resources/html/insert-logo.html"
    nature:
      beforeInit: ["./resources/js/macro.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: true
      ratio: "16:9" # If you change it to 4:3, open insert-logo.html to fix the watermark position.
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F,
                      dpi = 600,
                      fig.asp = 0.5, # You may change these parameters for each chunk as needed.
                      fig.align="center",
                      out.width = "80%")

library(tidyverse)



#For exporting as pdf, run 'pagedown::chrome_print('html')' in the console, where 'html' is the name of html file to be exported as pdf (in this case, 'template.html').

```

---
# Today's outline

1. A conceptual background to (computational) text analysis in R

2. __Basic text analysis tasks performed on vector strings following the logic of the tm package__

3. More advance text analysis tasks conducted on digitized books based on the tidytext package

4. Individually solving the problem set by building your own Gutenberg corpus
---
# Loading character vector documents (1/2)
To get going with our text analysis we need something to work with and to make it really simple we can generate our own documents by writing a few character vectors, such as a few juicy quotes from classical sociologists.
---

# Loading character vector documents (2/2)
```{r, echo = T, eval = F}
Soctxt.raw <- c(Durkheim = "Sociological method as we practice it rests wholly on the basic principle that social facts must be studied as things, that is, as realities external to the individual. There is no principle for which we have received more criticism; but none is more fundamental. Indubitably for sociology to be possible, it must above all have an object all its own. It must take cognizance of a reality which is not in the domain of other sciences... there can be no sociology unless societies exist, and that societies cannot exist if there are only individuals.", 
               Weber = "'Sociology' is a word which is used in many different senses. In the sense adopted here, it means the science whose object is to interpret the meaning of social action and thereby give a causal explanation of the way in which the action proceeds and the effects which it produces. By 'action' in this definition is meant human behaviour when and to the extent that the agent of agents see it as subjectively meaningful: the behaviour may be either internal or external, and may consist in the agent's doing something, omitting to do something, or having something done to him. By 'social' action is meant an action in which the meaning intended by the agent or agents involves a relation to another person's behaviour and in which that relation determines the way in which the action proceeds.", 
               Simmel = "I UNDERSTAND the task of sociology to be description and determination of the historico-psychological origin of those forms in which interactions take place between human beings. The totality of these interactions, springing from the most diverse impulses, directed toward the most diverse objects, and aiming at the most diverse ends, constitutes 'society'. Those different contents in connection with which the forms of interaction manifest themselves are the subject-matter of special sciences. These contents attain the character of social facts by virtue of occurring in this particular form in the interactions of men.",
              Tarde = "I will pass over a number of secondary objections which the application of the sociological point of view may encounter along its way. Since, after all, the fundamental nature of things is strictly inaccessible, and we are obliged to construct hypotheses in order to penetrate it, let us openly adopt this one and push it to its conclusion. Hypotheses fingo, I say naively. What is dangerous in the sciences are not tightly linked conjectures, logically followed to the ultimate depths or the ultimate precipices, but rather the ghosts of ideas which float aimlessly in the mind. The universal sociological point of view seems to me to be one of these spectres which haunt the brains of our speculative contemporaries.")

```

```{r}
Soctxt.raw <- c(Durkheim = "Sociological method as we practice it rests wholly on the basic principle that social facts must be studied as things, that is, as realities external to the individual. There is no principle for which we have received more criticism; but none is more fundamental. Indubitably for sociology to be possible, it must above all have an object all its own. It must take cognizance of a reality which is not in the domain of other sciences... there can be no sociology unless societies exist, and that societies cannot exist if there are only individuals.", 
               Weber = "'Sociology' is a word which is used in many different senses. In the sense adopted here, it means the science whose object is to interpret the meaning of social action and thereby give a causal explanation of the way in which the action proceeds and the effects which it produces. By 'action' in this definition is meant human behaviour when and to the extent that the agent of agents see it as subjectively meaningful: the behaviour may be either internal or external, and may consist in the agent's doing something, omitting to do something, or having something done to him. By 'social' action is meant an action in which the meaning intended by the agent or agents involves a relation to another person's behaviour and in which that relation determines the way in which the action proceeds.", 
               Simmel = "I UNDERSTAND the task of sociology to be description and determination of the historico-psychological origin of those forms in which interactions take place between human beings. The totality of these interactions, springing from the most diverse impulses, directed toward the most diverse objects, and aiming at the most diverse ends, constitutes 'society'. Those different contents in connection with which the forms of interaction manifest themselves are the subject-matter of special sciences. These contents attain the character of social facts by virtue of occurring in this particular form in the interactions of men.",
              Tarde = "I will pass over a number of secondary objections which the application of the sociological point of view may encounter along its way. Since, after all, the fundamental nature of things is strictly inaccessible, and we are obliged to construct hypotheses in order to penetrate it, let us openly adopt this one and push it to its conclusion. Hypotheses fingo, I say naively. What is dangerous in the sciences are not tightly linked conjectures, logically followed to the ultimate depths or the ultimate precipices, but rather the ghosts of ideas which float aimlessly in the mind. The universal sociological point of view seems to me to be one of these spectres which haunt the brains of our speculative contemporaries.")

```

---
# Creating a corpus (vector source)
To facilitate the analysis we can built a corpus object so that we can easily run each task on the whole body of texts.
```{r, echo = T, eval = F}
library(tm) # package for general text mining tasks
Soctxt.corp <- VCorpus(VectorSource(Soctxt.raw)) # create a volatile corpus, kept in memory as a R object.
inspect(Soctxt.corp) # let us have an overarching look with the tm function inspect
```

```{r}
library(tm) #Package for text mining tasks

Soctxt.corp <- VCorpus(VectorSource(Soctxt.raw)) # create a volatile corpus, kept in memory as a R object.

inspect(Soctxt.corp)
```

---
# Exporting corpus and importing txt
Perhaps we would at some point like to export our corpus from Rstudio to a folder on our disk. This can be made conveniently with the tm package. 
```{r, echo = T, eval = F}
writeCorpus(Soctxt.corp) # write corpus disk
Soctxt.dir <- VCorpus(DirSource(pattern = ".txt")) # generate a new corpus directly from disk; default is reading txt from the working directory
inspect(Soctxt.dir)
```

```{r}
writeCorpus(Soctxt.corp)

Soctxt.dir <- VCorpus(DirSource(pattern = ".txt")) #default is reading plain texts from working directory, but can be specified, see getReaders(), e.g, VCorpus(DirSource("/User/Directory", pattern = "Das_Kapital"), readerControl = list(reader=readDOC, language="German"))

inspect(Soctxt.dir)
```

---
# Managing corpus metadata
The two  main reasons for why you would like to have a corpus object is to store the documents' textual content and metadata. The later is use to keeping track of adn compare the documents. With the following code, we can couple the quote with its author. 
```{r, echo = T, eval = F}
meta(Soctxt.corp[[1]], "author") <- "Durkheim" # select the first quote and add the name to the author variable.
meta(Soctxt.corp[[2]], "author") <- "Weber"
meta(Soctxt.corp[[3]], "author") <- "Simmel"
meta(Soctxt.corp[[4]], "author") <- "Tarde"

meta(Soctxt.corp[[3]]) # take a look at the third quoute by Simmel
```

```{r}
meta(Soctxt.corp[[1]], "author") <- "Durkheim"
meta(Soctxt.corp[[2]], "author") <- "Weber"
meta(Soctxt.corp[[3]], "author") <- "Simmel"
meta(Soctxt.corp[[4]], "author") <- "Tarde"

meta(Soctxt.corp[[3]])
```

---
# Preprocessing - to lower case
We will here work with the bag-of-words model and, thus, we want words with the same meaning to end up in the same count. To enable this, we need to do some preprocessing tasks with the tm_map function. First off, we would like to treat words with uppercase and/or lowercase letters as the same. 
```{r, echo = T, eval = F}
Soctxt.corp.low <- tm_map(Soctxt.corp, content_transformer(tolower)) # we transform all letters to lowercase

Soctxt.corp.low[[3]]$content # What happened to Simmel?
```

```{r}
Soctxt.corp.low <- tm_map(Soctxt.corp, content_transformer(tolower))

Soctxt.corp.low[[3]]$content # What happened to Simmel?
```
---
# Preprocessing - remove punctuation
Second, punctuation within and around words can also disturb our preferred word couplings (e.g. "full-text" and "fulltext").
```{r, echo = T, eval = F}
Soctxt.corp.punct <- tm_map(Soctxt.corp, removePunctuation) # we simply remove all forms of puntuations

Soctxt.corp.punct[[3]]$content # What happened to Simmel?
```

```{r}
Soctxt.corp.punct <- tm_map(Soctxt.corp, removePunctuation)

Soctxt.corp.punct[[3]]$content # What happened to Simmel?
```

---
# Preprocessing - managing stop words (1/2)
Thirdly, too common and indistinct words, which are refered to as stop words, will create noise for most text analysis tasks. Thus, these will have to be removed. We can both make up our own set of words and rely on universal dictionaries of words, in this case English stop words.
```{r, echo = T, eval = F}
Soctxt.stop.words <- c("something", "can", "must", "since")

stopwords("en")
```

```{r}
Soctxt.stop.words <- c("something", "can", "must", "since")

stopwords("en")
```
---
# Preprocessing - managing stop words (2/2)
```{r, echo = T, eval = F}
Soctxt.corp.stop <- tm_map(Soctxt.corp, removeWords, stopwords("en"))
Soctxt.corp.stop <- tm_map(Soctxt.corp, removeWords, Soctxt.stop.words)

Soctxt.corp.stop[[3]]$content # What happened to Simmel?
```

```{r}
Soctxt.corp.stop <- tm_map(Soctxt.corp, removeWords, stopwords("en"))
Soctxt.corp.stop <- tm_map(Soctxt.corp, removeWords, Soctxt.stop.words)

Soctxt.corp.stop[[3]]$content # What happened to Simmel?
```
---

# Preprocessing - strip whitespace
Thirdly, white space 
```{r, echo = T, eval = F}
stripcorpus <- tm_map(Soctxt.corp, stripWhitespace)

stripcorpus[[3]]$content # What happened to Simmel?
```

```{r}
stripcorpus <- tm_map(Soctxt.corp, stripWhitespace)

stripcorpus[[3]]$content # What happened to Simmel?
```
---

# Preprocessing - stemming
```{r, echo = T, eval = F}
library(SnowballC) # Package for stemming
stemcorpus <- tm_map(Soctxt.corp, stemDocument)

stemcorpus[[3]]$content # What happened to Simmel?
```

```{r}
library(SnowballC) # Package for stemming
stemcorpus <- tm_map(Soctxt.corp, stemDocument)

stemcorpus[[3]]$content # What happened to Simmel?
```
---
# Applying all preprocessing tasks
```{r, echo = T, eval = F}
Soctxt.corp.clean <- Soctxt.corp %>%
  tm_map(content_transformer(tolower)) %>%
  tm_map(removePunctuation, preserve_intra_word_dashes = TRUE)  %>%
  tm_map(removeWords, stopwords("en")) %>%
  tm_map(stemDocument) %>%
  tm_map(stripWhitespace)

Soctxt.corp.clean[[3]]$content # What happened to Simmel?
```

```{r}
Soctxt.corp.clean <- Soctxt.corp %>%
  tm_map(content_transformer(tolower)) %>%
  tm_map(removePunctuation, preserve_intra_word_dashes = TRUE)  %>%
  tm_map(removeWords, stopwords("en")) %>%
  tm_map(stemDocument) %>%
  tm_map(stripWhitespace)

Soctxt.corp.clean[[3]]$content # What happened to Simmel?
```

---
# Generating a document-term matrix
```{r, echo = T, eval = F}
Soctxt.dtm <- DocumentTermMatrix(Soctxt.corp.clean) #Generate dtm from the preprocessed corpus

Soctxt.dtm.clean <- DocumentTermMatrix(Soctxt.corp.clean, #Generate and preprocessing a dtm from the original corpus
                               control = list(removePunctuation = TRUE, stripWhitespace = TRUE,
                                              removeSparseTerms = 0.99, stopwords = TRUE,
                                              stemming = TRUE))
inspect(Soctxt.dtm) #Inspect dtm
```

```{r}
Soctxt.dtm <- DocumentTermMatrix(Soctxt.corp.clean) #Generate dtm from the preprocessed corpus

Soctxt.dtm.clean <- DocumentTermMatrix(Soctxt.corp.clean, #Generate and preprocessing a dtm from the original corpus
                               control = list(removePunctuation = TRUE, 
                                              stripWhitespace = TRUE,
                                              removeSparseTerms = 0.99, 
                                              stopwords = TRUE,
                                              stemming = TRUE))
inspect(Soctxt.dtm) #Inspect dtm
```
---
# Operating a dtm - terms per corpus
```{r, echo = T, eval = F}
findFreqTerms(Soctxt.dtm, 4) # find terms appearing at least 4 times
```

```{r}
findFreqTerms(Soctxt.dtm, 4) # find terms appearing at least 4 times
```
---
# Operating a dtm - terms per doc
```{r, echo = T, eval = F}
findMostFreqTerms(Soctxt.dtm) # find most frequent terms for each document
```

```{r}
findMostFreqTerms(Soctxt.dtm)
```

---
# Operating a dtm - term correlations
```{r, echo = T, eval = F}
findAssocs(Soctxt.dtm, terms = "sociolog", corlimit = 0.6) # terms correlating to a specific word
```

```{r}
findAssocs(Soctxt.dtm, terms = "sociolog", corlimit = 0.6) # terms correlating to a specific word
```
---
# Creating dicionaries
```{r, echo = T, eval = F}
inspect(DocumentTermMatrix(cleancorpus, list(dictionary=c("sociolog","interact","social")))) #dictionary
```

```{r}
inspect(DocumentTermMatrix(cleancorpus, list(dictionary=c("sociolog","interact","social")))) #dictionary
```
---
# BONUS: From dtm to word cloud
```{r, echo = T, eval = F}
library(wordcloud) #Packge for generating wordclouds
Soctxt.df <- as.matrix(Soctxt.dtm)
Soctxt.df <- sort(colSums(Soctxt.df),decreasing=TRUE)
Soctxt.df <- data.frame(word = names(Soctxt.df),freq=Soctxt.df)

wordcloud(df$word,df$freq,colors = brewer.pal(12, "Dark2"))
```
---
# BONUS: From dtm to word cloud (viz)
```{r}
library(wordcloud) #Packge for generating wordclouds
Soctxt.df <- as.matrix(Soctxt.dtm)
Soctxt.df <- sort(colSums(Soctxt.df),decreasing=TRUE)
Soctxt.df <- data.frame(word = names(Soctxt.df),freq=Soctxt.df)

wordcloud(df$word,df$freq,colors = brewer.pal(12, "Dark2"))
```

---
# Thank you for your time!

## Do not hesitate to contact me

|                                                                                              |                          |
|:---------------------------------------------------------------------------------------------|:-------------------------|
| <a href="mailto:josef.ginnerskov@soc.uu.se">.UUred[<i class="fa fa-paper-plane fa-fw"></i>] |josef.ginnerskov@soc.uu.se |  
| <a href="http://twitter.com/doeparen">.UUred[<i class="fa fa-twitter fa-fw"></i>]         |@doeparen              |
| <a href="http://github.com/doeparen">.UUred[<i class="fa fa-gitlab fa-fw"></i>]           |@doeparen              |