---
title: "Text analysis II: Building a tidytext toolbox"
subtitle: "Introduction to R for Social Sciences (Sociology ∞ HERO)"
author: 
- "Josef Ginnerskov, doctoral candidate"
# Uncomment the line below for adding more authors
#- "Another Author"
institute: "Department of Sociology"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    self_contained: true # This allows to use the html file without copying all files.
    lib_dir: libs
    css: ["./resources/css/uppsala.css", "./resources/css/uppsala-fonts.css"]
    includes:
      in_header: "./libs/partials/header.html"
#      after_body: "./resources/html/insert-logo.html"
    nature:
      beforeInit: ["./resources/js/macro.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: true
      ratio: "16:9" # If you change it to 4:3, open insert-logo.html to fix the watermark position.
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F,
                      dpi = 600,
                      fig.asp = 0.5, # You may change these parameters for each chunk as needed.
                      fig.align="center",
                      out.width = "80%")

library(tidyverse)

#For exporting as pdf, run 'pagedown::chrome_print('html')' in the console, where 'html' is the name of html file to be exported as pdf (in this case, 'template.html').

```

# Loading books from www.gutenberg.org
```{r, echo = T, eval = F}
library(gutenbergr)

sociology_raw <- gutenberg_download(c(41360, 46423, 13205, 30610, 6568, 21609), 
                                    meta_fields = c("author", "title"), strip = TRUE)
head(sociology_raw)
```

```{r}
library(gutenbergr)

sociology_raw <- gutenberg_download(c(41360, 46423, 13205, 30610, 6568, 21609), # Ellwood, Durkheim, Marx
                                    meta_fields = c("author", "title"), strip = TRUE)
head(sociology_raw)
```
---
# Preprocessing - fix awkward signs
```{r, echo = T, eval = F}
sociology_raw$text <- gsub("_","",as.character(sociology_raw$text))
sociology_raw$text <- gsub("--"," ",as.character(sociology_raw$text))
```

```{r}
sociology_raw$text <- gsub("--"," ",as.character(sociology_raw$text))
sociology_raw$text <- gsub("_","",as.character(sociology_raw$text))
```
---

# Turning books into a tidy tibble
```{r, echo = T, eval = F}
library(tidyverse)
library(tidytext)
tidy_sociology <- sociology_raw %>%
    mutate(line = row_number()) %>%
    unnest_tokens(word, text)

head(tidy_sociology)
```

```{r}
library(tidyverse)
library(tidytext)
tidy_sociology <- sociology_raw %>%
    mutate(line = row_number()) %>%
    unnest_tokens(word, text)

head(tidy_sociology)
```
---
# Removing stop words
```{r, echo = T, eval = F}
tidy_sociology <- tidy_sociology %>%
  anti_join(stop_words)

my_words <- tibble(word = c("p", "cit", "tr", "pp", "ff", "nat", "ibid", "geddes", "prof", "per"))

tidy_sociology <- tidy_sociology %>%
  anti_join(my_words)
```

```{r}
tidy_sociology <- tidy_sociology %>%
  anti_join(stop_words)

my_words <- tibble(word = c("p", "cit", "tr", "pp", "ff", "nat", "ibid", "geddes", "prof", "per"))

tidy_sociology <- tidy_sociology %>%
  anti_join(my_words)
```

---
# Basic global/corpus count
```{r, echo = T, eval = F}
tidy_sociology %>%
    count(word, sort = TRUE) %>%
    top_n(10)
```

```{r}
tidy_sociology %>%
    count(word, sort = TRUE) %>%
    top_n(10)
```

---
# Basic local/document count
```{r, echo = T, eval = F}
tidy_sociology %>%
    filter(author == "Marx, Karl") %>%
    count(word, sort = TRUE) %>%
    top_n(10)
```

```{r}
tidy_sociology %>%
    filter(author == "Marx, Karl") %>%
    count(word, sort = TRUE) %>%
    top_n(10)
```

---
# Generating a word cloud (a)

```{r, echo = T, eval = F}
library(wordcloud)

tidy_sociology %>%
  count(word) %>%
  with(wordcloud(word, n, 
                 max.words = 100, 
                 colors = brewer.pal(8, "Dark2")))
```
---
# Generating a word cloud (b)
```{r}
library(wordcloud)

tidy_sociology %>%
  count(word) %>%
  with(wordcloud(word, n, 
                 max.words = 100, 
                 colors = brewer.pal(8, "Dark2")))
```

---
# Comparing word occurences per book
```{r, echo = T, eval = F}
words_sociology <- tidy_sociology %>%
  count(author, word, sort = TRUE)

words_sociology_work <- words_sociology %>% 
  group_by(author) %>% 
  summarize(total = sum(n))

words_sociology_work <- left_join(words_sociology_work, words_sociology)
```

```{r}
words_sociology <- tidy_sociology %>%
  count(author, word, sort = TRUE)

words_sociology_work <- words_sociology %>% 
  group_by(author) %>% 
  summarize(total = sum(n))

words_sociology_work <- left_join(words_sociology, words_sociology_work)

words_sociology_work
```
---
# Viz word occurences per book (a)
```{r, echo = T, eval = F}
words_sociology_work %>% filter (n >= 333) %>% 
   mutate(word = reorder(word, n)) %>% 
   ggplot(aes(x = word, y = n, fill = author)) + 
   geom_col() + 
   coord_flip()
```
---
# Viz word occurences per book (b)

```{r}
words_sociology_work %>% filter (n >= 333) %>% 
   mutate(word = reorder(word, n)) %>% 
   ggplot(aes(x = word, y = n, fill = author)) + 
   geom_col() + 
   coord_flip()
```
---
# Comparing books via td-idf

```{r, echo = T, eval = F}
sociology_tf_idf <- tidy_sociology %>%
    count(author, word, sort = TRUE) %>%
    bind_tf_idf(word, author, n) %>%
    arrange(-tf_idf) %>%
    group_by(author) %>%
    top_n(15) %>%
    ungroup
```

```{r}
sociology_tf_idf <- tidy_sociology %>%
    count(author, word, sort = TRUE) %>%
    bind_tf_idf(word, author, n) %>%
    arrange(-tf_idf) %>%
    group_by(author) %>%
    top_n(15) %>%
    ungroup
sociology_tf_idf
```
---
# Viz books via td-idf (a)

```{r, echo = T, eval = F}
sociology_tf_idf %>%
    mutate(word = reorder_within(word, tf_idf, author)) %>%
    ggplot(aes(word, tf_idf, fill = author)) +
    geom_col(alpha = 0.8, show.legend = FALSE) +
    facet_wrap(~ author, scales = "free", ncol = 3) +
    scale_x_reordered() +
    coord_flip() +
    theme(strip.text=element_text(size=11))
```
---
# Viz books via td-idf (b)

```{r}
sociology_tf_idf %>%
    mutate(word = reorder_within(word, tf_idf, author)) %>%
    ggplot(aes(word, tf_idf, fill = author)) +
    geom_col(alpha = 0.8, show.legend = FALSE) +
    facet_wrap(~ author, scales = "free", ncol = 3) +
    scale_x_reordered() +
    coord_flip() +
    theme(strip.text=element_text(size=11))
```

---
# Calculating word correlations

```{r, echo = T, eval = F}
library(widyr)
cors_sociology <- words_sociology_work %>% 
  group_by(word) %>%
  filter(n >= 250) %>%
  pairwise_cor(word, author, sort = TRUE, upper = FALSE)
```

```{r}
library(widyr)
cors_sociology <- words_sociology_work %>% 
  group_by(word) %>%
  filter(n >= 250) %>%
  pairwise_cor(word, author, sort = TRUE, upper = FALSE)

cors_sociology
```
---

# Word correlations -> word graph
```{r, echo = T, eval = F}
library(igraph)
set.seed(1234)
g_keywords <- cors_sociology %>%
  filter(correlation > .3) %>%
  graph_from_data_frame()
```

```{r}
library(igraph)
set.seed(1234)
g_keywords <- cors_sociology %>%
  filter(correlation > .3) %>%
  graph_from_data_frame()
g_keywords
```

---
# Viz the word graph (a)
```{r, echo = T, eval = F}
library(ggraph)
g_keywords %>% 
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation, edge_width = correlation), edge_colour = "orange") +
  geom_node_point(size = 0.5*igraph::degree(g_keywords), colour = "lightblue") +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_void()
```

---
# Viz the word graph (b)
```{r}
library(ggraph)
g_keywords %>% 
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation, edge_width = correlation), edge_colour = "orange") +
  geom_node_point(size = 0.5*igraph::degree(g_keywords), colour = "lightblue") +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_void()
```

---
# Preparing for topic modeling

```{r, echo = T, eval = F}
sociology_dtm <- words_sociology_work %>%
  cast_dtm(author, word, n)
```

```{r}
sociology_dtm <- words_sociology_work %>%
  cast_dtm(author, word, n)

sociology_dtm
```
---
# Run a topic model
```{r, echo = T, eval = F}
library(topicmodels)
sociology_lda <- LDA(sociology_dtm, k = 4, control = list(seed = 1234))
```

```{r}
library(topicmodels)
sociology_lda <- LDA(sociology_dtm, k = 4, control = list(seed = 1234))
sociology_lda
```
---
# Inspecting the beta (word/topic prob)

```{r, echo = T, eval = F}
sociology_topics <- tidy(sociology_lda, matrix = "beta")
```

```{r}
sociology_topics <- tidy(sociology_lda, matrix = "beta")
sociology_topics
```
---
# Top terms in topic model beta

```{r, echo = T, eval = F}
top_terms <- sociology_topics %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>% 
  ungroup() %>%
  arrange(topic, -beta)
```

```{r}
top_terms <- sociology_topics %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>% 
  ungroup() %>%
  arrange(topic, -beta)
top_terms
```
---
# Viz topic model beta (a)

```{r, echo = T, eval = F}
library(ggplot2)

top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  scale_y_reordered()
```
---
# Viz topic model beta (b)

```{r}
library(ggplot2)

top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  scale_y_reordered()
```

---
# Inspecting the gamma (doc/topic prob)
```{r, echo = T, eval = F}
sociology_gamma <- tidy(sociology_lda, matrix = "gamma")
```

```{r}
sociology_gamma <- tidy(sociology_lda, matrix = "gamma")
sociology_gamma
```
---
# Viz topic model gamma (a)

```{r, echo = T, eval = F}
sociology_gamma %>%
  mutate(document = reorder(document, gamma * topic)) %>%
  ggplot(aes(factor(topic), gamma)) +
  geom_boxplot() +
  facet_wrap(~ document) +
  labs(x = "topic", y = expression(gamma))
```
---
# Viz topic model gamma (b)

```{r}
sociology_gamma %>%
  mutate(document = reorder(document, gamma * topic)) %>%
  ggplot(aes(factor(topic), gamma)) +
  geom_boxplot() +
  facet_wrap(~ document) +
  labs(x = "topic", y = expression(gamma))
```
---
# Adding sentiments to your tibble
```{r, echo = T, eval = F}
bing <- get_sentiments("bing")

tidy_sociology %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE) %>%
  top_n(10)
```

```{r}
bing <- get_sentiments("bing")

tidy_sociology %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE) %>%
  top_n(10)
```
---
# Viz top sentiments (a)
```{r, echo = T, eval = F}
sentiment_counts <- tidy_sociology %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE)

sentiment_counts %>%
  filter(n > 100) %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col() +
  coord_flip() +
  labs(y = "Contribution to sentiment")
```
---
# Viz top sentiments (b)
```{r}
sentiment_counts <- tidy_sociology %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE)


sentiment_counts %>%
  filter(n > 100) %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col() +
  coord_flip() +
  labs(y = "Contribution to sentiment")
```
---
# The sentiment narrative unfolds (a)

```{r, echo = T, eval = F}
library(tidyr)

sentisociology <- tidy_sociology %>%
  inner_join(bing) %>%
  count(author, index = line %/% 100, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

```{r}
library(tidyr)

sentisociology <- tidy_sociology %>%
  inner_join(bing) %>%
  count(author, index = line %/% 100, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

```{r, echo = T, eval = F}
library(ggplot2)

ggplot(sentisociology, aes(index, sentiment, fill = author)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~author, ncol = 2, scales = "free_x")
```
---
# The sentiment narrative unfolds (b)

```{r}
library(ggplot2)

ggplot(sentisociology, aes(index, sentiment, fill = author)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~author, ncol = 2, scales = "free_x")
```


---

# Thank you for your time!

## Do not hesitate to contact me

|                                                                                              |                          |
|:---------------------------------------------------------------------------------------------|:-------------------------|
| <a href="mailto:josef.ginnerskov@soc.uu.se">.UUred[<i class="fa fa-paper-plane fa-fw"></i>] |josef.ginnerskov@soc.uu.se |  
| <a href="http://twitter.com/doeparen">.UUred[<i class="fa fa-twitter fa-fw"></i>]         |@doeparen              |
| <a href="http://github.com/doeparen">.UUred[<i class="fa fa-gitlab fa-fw"></i>]           |@doeparen              |